---
title: "PA4"
subtitle: "Data Science for Linguists"
author: "Kaylee Fernandez"
date: "3/30/2025"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---
# load libraries

```{r}
library("dplyr")
library("readr")
library("tidyverse")
library("ds4ling")
library("patchwork")
library("ggplot2")
library("lmtest")
```

---

# load raw data

```{r}
fricatives_raw_data <- read_csv("../data_raw/fricatives/fricatives.csv")
```

---

# inspect data

```{r}
str(fricatives_raw_data)
```

---
# tidy data

```{r}
fricatives_tidy <- fricatives_raw_data |>
  pivot_longer( #change data from wide to long format
    cols = c(s_cog, s_skewness, sh_cog, sh_skewness), #old columns, 
    names_to = c("sound", ".value"), #separate each column into sound and variable
    names_sep = "_" #split the name at the underscore
  )

str(fricatives_tidy)
write_csv(fricatives_tidy, "../data_tidy/fricatives_tidy.csv")
```

---

## boxpot of center of gravity as a function of phoneme
```{r, echo=FALSE}
fricatives_tidy |>
  ggplot() +
  aes(x = factor(sound), y = cog) + #factor tells R to treat sound as a categorical value
  geom_boxplot()
```


---
## skewness as a function of phoneme

```{r, echo=FALSE}
fricatives_tidy |>
  ggplot() + # you add + when you are adding a new layer 
  aes(x = factor(sound), y = skewness) +
  stat_summary(
    fun.data = mean_sdl, # plot mean and standard deviation
    geom = "pointrange"  
  )
```

---
### model examining center of gravity as a function of skewness for the [s] segments
```{r, echo=FALSE}
s_data <- fricatives_tidy |>
  filter(sound == "s")

s_model <- lm(cog ~ skewness, data = s_data)
summary(s_model)
```
---
# scatterplot of model: figure 1
```{r, echo=FALSE, message=FALSE}
s_model |>
  ggplot() +
  aes(x = skewness, y = cog) +
  geom_point() + # add geom scatter plot
  geom_smooth(method = "lm") + # add regression line
  labs(
    title = "Center of gravity as a function of skewness for [s]",
    x = "Skewness",
    y = "Center of gravity (Hz)"
  )
```

---
class: center, middle

#check model diagnostics (assumptions)

---
### 1: The regression model is linear in parameters -- *yes!*

```{r, echo=FALSE, message=FALSE}
s_model |>
  ggplot() +
  aes(x = skewness, y = cog) +
  geom_point() # add geom scatter plot
```

---

### 2: The mean of residuals is zero -- *yes, close to zero!*

```{r, echo=FALSE, message=FALSE}
options(scipen = 999)
mean(s_model$residuals)
```


---

### 3: Homoskedasticity of residuals or equal variance --- *yes, spread of residuals is the same above and below the line!*

```{r, echo=FALSE, message=FALSE}
diagnosis(s_model)
```

---

### 4: No autocorrelation of residuals: *Durbin-Watson test is not significant, no autocorrelation!*

```{r}
dwtest(s_model)
```

---

### 5: Predictor and residuals are not correlated: *confirmed, correlation is near 0 so we know predictor and residuals are not correlated!*

```{r}
cor.test(s_data$skewness, residuals(s_model))
```

---

### 6: Normality of residuals: *confirmed, residuals look approximately normally distributed*

```{r, echo=FALSE, message=FALSE}
diagnosis(s_model)
summary(s_model)
```

---

### 7: General description of results

#### 
Data from fricatives were analyzed using a linear model. The model included center of gravity measurement in Hz for [s] as the dependent variable and skewness measurement for [s] as the independent variable. Visual inspection of Q-Q plots and plots of residuals against fitted values revealed no discernible deviations from normality and homoskedasticity. Additional diagnostics confirmed linearity, a near-zero mean of residuals, no autocorrelation, and no correlation between the predictor and residuals. We report p-values with alpha set as 0.05.

Figure 1 (see slide 9) shows center of gravity as a function of skewness for [s]. 

The model revealed a significant fixed effect of skewness . Specifically, higher skewness was associated with higher center of gravity values of [s] (*B* = 1675.60, SE = 139.40, *p* < .001). The model accounted for approximately 75% of the variance in center of gravity (R² = 0.75).

---

# assumptions dataset

```{r, echo=FALSE}
assumptions_data <- read_csv("../data_raw/fricatives/assumptions.csv")
str(assumptions_data)
```

---

```{r, echo=FALSE}
assumptions_data |>
  ggplot() +
  aes(x = x, y = y) +
  geom_point()
```

### The relationship between x and y is definitely not linear. It is actually a dinosaur relationship! We cannot fit a linear model to this dataset. 






